---
title: "Initial Plots"
author: "Jackson Dial"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(pander)
library(mosaic)
library(DT)
```

## Sources

[link](https://www.idaho-demographics.com/counties_by_population)

## Read in the Data

```{r message=FALSE, warning=FALSE}
data1 <- read_xlsx("Public Records Request 11-17-2020.xlsx")
data1$County <- factor(data1$County)

#Create a population-adjusted case count per county

data2 <- data1 %>% 
  mutate(dummy_var = 1) %>% 
  group_by(`Report Date`, County) %>%
  summarise(count_cases = sum(dummy_var)) %>% 
  mutate(population = case_when(
    County == "Bonneville" ~ 119062,
    County == "Clark" ~ 845,
    County == "Custer" ~ 4315,
    County == "Fremont" ~ 13099,
    County == "Jefferson" ~ 29871,
    County == "Lemhi" ~ 8027,
    County == "Madison" ~ 39907,
    County == "Teton" ~ 12142
  )) %>% 
  mutate(case_adj = (count_cases / population) * 10000) 



```

## Let's try some ggplots

```{r message=FALSE, warning=FALSE}
ggplot(data2, aes(x = `Report Date` ,y = count_cases, color = County))+
  geom_line()+
  facet_wrap(vars(County), ncol = 4)+
  theme_bw()+
  theme(legend.position = "none")


ggplot(data2, aes(x = `Report Date`, y = case_adj, color = County))+
  geom_line()+
  facet_wrap(vars(County), ncol = 4)+
  theme_bw()+
  theme(legend.position = "none")+
  labs(y = "New Cases per 10,000",
       title = "Population Adjusted Incidence Rates per County")

```

## Wide Data

```{r}
wide_data_cases <- pivot_wider(data2,
                         names_from = County,
                         values_from = count_cases,
                         values_fill = 0)

pairs(wide_data_cases)

wide_data_adj <- pivot_wider(data2,
                         names_from = County,
                         values_from = case_adj,
                         values_fill = 0)

pairs(wide_data_adj)

running_total_cases <- data2 %>% 
  group_by(County, `Report Date`) %>% 
  mutate(cum_sum_ever = cumsum(count_cases))

wide_running <- running_total_cases %>% pivot_wider(
  names_from = County,
  values_from = cum_sum_ever,
  values_fill = 0
)

pairs(wide_running)

```

## Duration Comparisons

```{r}
data_rename <- data1 %>% 
  rename(
    report_date =`Report Date`,
    released_date = `Released from EIPH Monitoring`,
    age_group = `Age Group (by decade)`,
    exposure = `Source of Exposure`
  )

# Compute the amount of time spent under EIPH monitoring
# Some values were negative, but it was assumed the dates were reversed by accident.
# This was accounted for by square-rooting the square of each value.

duration_monitored <- data_rename %>% 
  mutate(total_time = difftime(released_date, report_date, unit = "days")) %>% 
  mutate(pos_time = sqrt(as.numeric(total_time)^2))

ggplot(duration_monitored,aes(x = as.factor(County), y = pos_time))+
  geom_boxplot(aes(fill = County))+
  theme_bw()+
  theme(panel.grid.major.x = element_blank())

```

## ANOVA Test

```{r message=FALSE, warning=FALSE}
favstats(total_time~County, data = duration_monitored)

case.aov <- aov(pos_time ~ County, data = duration_monitored,
                contrasts = list(County = contr.sum))

pander(summary(case.aov))

```


## Pairwise Comparisons

```{r}
pairwise.t.test(duration_monitored$pos_time, duration_monitored$County, "none") %>% pander()
```

## Effects

```{r}
dummy.coef(case.aov) %>% pander()

```

## Wilcoxon Rank-Sum Test

```{r}
kruskal.test(pos_time ~ County, data = duration_monitored)

```

## Exposure vs County?

```{r}
#joined_data <- data1 %>% 
#  inner_join(data2, by = c(`Report Date` = `Report Date`, County = County))

exp_county <- data_rename %>% 
  group_by(exposure, County) %>% 
  summarise(count = length(ID)) %>% 
  mutate(exp = case_when(
    exposure %in% c("community Transmission", "Community Transmission", "Community TRansmission") ~ "Community Transmission",
    exposure %in% c("contact with a Confirmed Case", "Contact with a Confirmed Case", "Contact With a Confirmed Case") ~ "Contact with a confirmed case",
    exposure %in% c("Ongoing Investigation", "ONgoing Investigation") ~ "Ongoing investigation",
    exposure == "Travel" ~ "Travel"
  )) %>% 
  subset(select = -c(exposure))

ggplot(exp_county, aes(x = County, y = count, fill = exp))+
  geom_bar(stat = "identity", position = "dodge")+
  theme_bw()+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(y = "Number of Reported Cases",
       fill = "Exposure Source")

exp_wide <- exp_county %>% pivot_wider(names_from = County,
                                       values_from = count)

exp_agg <- exp_county %>% 
  group_by(County, exp) %>% 
  summarise(count_tot = sum(count))

exp_wide <- exp_agg %>% pivot_wider(names_from = County,
                                    values_from = count_tot,
                                    values_fill = 0)

#exp_count_table <- table(exp_wide$County, exp_wide$exp)

#pander(exp_count_table)
my_table <- as.matrix(exp_wide[,-1])

rownames(my_table) <- exp_wide[,1][[1]]


```

## Chi-Squared Test

```{r}
chisq.test(my_table[-4,])$expected %>% pander()
```


## Hospitalization and length of Supervision? Logistic Regression Model

```{r}
len_hosp <- data_rename %>% 
  mutate(length_hosp = difftime(released_date, report_date, unit = "days")) %>% 
  mutate(pos_len = sqrt(as.numeric(length_hosp)^2))


len_glm <- glm(formula = (Hospitalized == "Yes") ~ pos_len, data = len_hosp, family = "binomial")
summary(len_glm)

b <- coef(len_glm)

plot(formula = (Hospitalized == "Yes") ~ pos_len, data = len_hosp, pch = 16, col = rgb(.1,.1,.1,.05), xlab = "Length of Monitoring", ylab = "Probability of Being Hospitalized (1 = Hospitalized)")

curve(exp(b[1] + b[2]*x)/(1+exp(b[1]+b[2]*x)), add = TRUE)
library(ResourceSelection)
hoslem.test(len_glm$y, len_glm$fitted, g = 30)
```

## Hospitalization and Length

```{r}
sex_length_glm <- glm(formula = (Sex == "F") ~ pos_len, data = len_hosp, family = "binomial")
summary(sex_length_glm)
#insignificant p-values, no need to plot
```

## Length and Age Group

```{r}
table(len_hosp$age_group)
# Create a column classifying those form 0-39 as 0 and those from 40-90 as 1
two_age_groups <- len_hosp %>% 
  mutate(age_group_2 = case_when(
    age_group %in% c("0-4", "5-12", "13-17", "18-29", "30-39") ~ 0,
    TRUE ~ 1
  ))

age_glm <- glm(formula = (age_group_2 == 1) ~ pos_len, data = two_age_groups, family = "binomial")

summary(age_glm)

b1 <- coef(age_glm)

plot(formula = (age_group_2 == 1) ~ pos_len, data = two_age_groups, pch = 16, col = as.factor(age_group_2), xlab = "Length of Monitoring", ylab = "Probability of Being above 40")

curve(exp(b1[1] + b1[2]*x)/(1+exp(b1[1]+b1[2]*x)), add = TRUE)

```


```{r}
len_hosp <- len_hosp %>% 
  mutate(age_group = factor(age_group, levels = c("0-4", "5-12", "13-17", "18-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+"))) %>% 
  mutate(Hospitalized = case_when(
    Hospitalized == "Yes" ~ "Yes",
    TRUE ~ "No"
  ))
```

```{r}
ggplot(len_hosp, aes(x = as.factor(age_group), y = pos_len, fill = as.factor(Hospitalized)))+
  geom_bar(position = "stack", stat = "identity")+
  facet_wrap(vars(County), scales = "free")+
  theme(axis.text.x = element_text(angle = 30))+
  coord_flip()

#library(rnoaa)

#vignette("rnoaa")

#options(noaakey = "ETJJubDnGOQeIsueSktTGtoOAJBssfnw")

#weather_data <- ncdc_locs(locationcategoryid = "CITY", sortfield = "name", sortorder = "desc")

#eiph_cities <- ncdc_locs(locationcategoryid = "CITY", sortfield = "name", sortorder = "desc") %>% 
#  filter(name == "Rexburg, ID US")
#str(weather_data)

#grep("ID", weather_data$data$name, value = TRUE)

```

```{r}
library("devtools")
install_github("Ram-N/weatherData")
library(weatherData)
weather_data <- checkDataAvailabilityForDateRange("KIDA", "2020-01-01", "2020-11-20")

getDetailedWeather("KIDA", "2020-01-01", "id", 
        opt_temperature_columns = T, opt_compress_output = T, 
        opt_verbose = T)

```


### Geofacet per county of the progression of disease prevalence

```{r}
#library(geofacet)
#ggplot(data2, aes(x = `Report Date`, y = case_adj))+
#  facet_geo(vars(County))


```
